\thetitle

### 2021-05-14 writing/discussion.tex
trail examples: touch up figure and new prose

Interpreting the numeric summaries and aggregations of the last section demands an intuitive understanding of what blame trails look like in practice.
Thus, before drawing high level conclusions, let us examine one set of scenarios with a typical debugging scenario from the {\tt take5} benchmark,
whose module dependency graph is shown in the top left of figure~\ref{fig:example-trails}.

The rest of figure~\ref{fig:example-trails} illustrates the blame trails produced by every mode of the rational programmer for the debugging scenario in
two different ways: 
\begin{itemize}

\item The top right shows the blame trails produced by every mode of the rational programmer as paths through the configuration lattice starting at the
root (leftmost) configuration. Each configuration is represented by a bit string where each digit corresponds to a module in the program and indicates if
it is typed.

@@

Let us contrast rows 1 and~4. The first one shows that running the root
configuration under the Natural-blame mode fails due a dynamic type check and
blames the {\tt player} module; typing that module and running again then
results in a type error, and hence the trail is successful.  By contrast, the
Natural-exceptions mode (row 4) yields stack information for the root
configuration that is unhelpful; it identifies only {\tt main}, which is already
typed. Hence, this trail immediately gets stuck.

In short, this figure concretely demonstrates how the rational programmer
behaves in different modes. In this case, the behaviors differ from each other
in all five modes. The reader may keep these scenarios in mind as we next
discuss the numeric results of the last section.

>>>>

Interpreting the numeric summaries and aggregations of the last section demands an intuitive understanding of what blame trails look like in practice.
Thus, before drawing high level conclusions, let us examine one set of trails from a typical debugging scenario from the {\tt take5} benchmark,
whose module dependency graph is shown in the top left of figure~\ref{fig:example-trails}.

The rest of figure~\ref{fig:example-trails} illustrates the blame trails produced by every mode of the rational programmer (except Random) for the debugging scenario in
two different ways: 
\begin{itemize}

\item The top right shows the blame trails produced by every mode of the rational programmer as paths through the configuration lattice starting at the
root (leftmost) configuration. Each configuration is represented by a sequence of boxes corresponding to modules in the program, with an x indicating that the module is typed.
The mutated module has a gray box.

@@

In illustration of how to read the table, let us contrast rows 1 and~4. The first one shows that running the root
configuration under the Natural-blame mode fails due a dynamic type check and
blames the {\tt player} module; typing that module and running again then
results in a type error, and hence the trail is successful.  By contrast, the
Natural-exceptions mode (row 4) yields stack information for the root
configuration that is unhelpful; it identifies only {\tt main}, which is already
typed. Hence, this trail immediately gets stuck.

In short, this figure concretely demonstrates how the rational programmer
behaves in different modes. In this case, the behaviors differ from each other
in five of the six modes (the two Transient blame modes behave the same).
The reader may keep these scenarios in mind as we next
discuss the numeric results of the last section.

### 2021-05-14 writing/trail-example-table.tex
trail examples: touch up figure and new prose

\noindent{\bf result}\\
\begin{center}
\begin{tabular}{l@{\quad}l}
symbol        & denotation \\ \hline 
\blameFinger  & the configuration results in a dynamic type-check failure, blaming the module(s) below \\
\typeError    & the configuration does not type check\\
\runtimeError & the configuration fails a check to the runtime system\\
\checkFailure & the configuration signals a dynamic type check failure for which blame is ignored\\
\end{tabular}
\end{center}

>>>>

\noindent{\bf result}\\
\begin{center}
\vspace{-2.5em}
\begin{tabular}{l@{\quad}l}
symbol        & denotation \\ \hline 
\blameFinger  & the configuration signals a dynamic type check failure, blaming the module(s) below \\
\typeError    & the configuration does not type check\\
\runtimeError & the configuration fails a check by the runtime system\\
\checkFailure & the configuration signals a dynamic type check failure for which blame is ignored\\
\end{tabular}
\end{center}
